Chord: A Versatile Platform for Program Analysis

### 2011-03-10 main/doc/deadlock.tex
added more documentation

\section{Deadlock analysis}
\label{sec:deadlock}

Run the deadlock analysis provided in Chord by running the following
command in Chord's {\tt main/} directory:

@@

\noindent where {\tt <...>} is the absolute path of a directory
containing a file named \code{chord.properties} which defines
properties \code{chord.main.class}, \code{chord.class.path}, and
\code{chord.src.path}.  See Section \ref{sec:chord-sysprops} for the
meaning of these properties.

Directory \code{main/examples/deadlock_test/} provides a toy Java
program on which you can run the deadlock analysis.  First run {\tt
  ant} in that directory (in order to compile the program's {\tt
  .java} files to {\tt .class} files) and then run the above command
in Chord's {\tt main/} directory with {\tt <...>} replaced by
\code{/.../examples/deadlock_test/}.  Upon successful completion, the
file \code{deadlocks.html} should be produced in directory
\code{main/examples/deadlock_test/chord_output/}.

>>>>

\section{Deadlock Analysis}
\label{sec:deadlock}

Run Chord's static deadlock analysis by running the following
command in Chord's {\tt main/} directory:

@@

\noindent where {\tt <...>} denotes the absolute or relative path of a directory
containing a file named \code{chord.properties} which defines
properties \code{chord.main.class}, \code{chord.class.path}, and
\code{chord.src.path}.  See Section \ref{sec:chord-sysprops} for the
meaning of these properties.

Directory \code{main/examples/deadlock_test/} provides a toy Java
program on which you can run the deadlock analysis.  First run {\tt
  ant} in that directory (in order to compile the program's {\tt
  .java} files to {\tt .class} files) and then run the above command
in Chord's {\tt main/} directory with {\tt <...>} replaced by
\code{examples/deadlock_test/}.  Upon successful completion, the
file \code{deadlocks.html} should be produced in directory
\code{examples/deadlock_test/chord_output/}.

### 2011-03-10 main/doc/download_and_installation.tex
added more documentation

Ensure that the following software is installed on your machine:

\begin{itemize}
\item
JDK 6 or higher, e.g. from \htmladdnormallink{IBM}{http://www.ibm.com/developerworks/java/jdk/} or
\htmladdnormallink{Sun}{http://java.sun.com/javase/downloads/index.jsp}
\item
\htmladdnormallink{Apache Ant}{http://ant.apache.org/}, a Java-based build tool
\item
a C++ compiler, e.g. \htmladdnormallink{GCC}{http://gcc.gnu.org/}
\item
a Make utility, e.g. \htmladdnormallink{GNU Make}{http://www.gnu.org/software/make/}
\item
\htmladdnormallink{Cygwin}{http://www.cygwin.com/}, if it is a Windows machine
\end{itemize}

Download a \Chord\ source release from \url{http://code.google.com/p/jchord/downloads/list} or
the latest source from the SVN repository at \url{http://code.google.com/p/jchord/source/checkout}.

Directory {\tt main/} contains a {\tt build.xml} file which is interpreted by Apache Ant.
To see the various targets available, run the following command in that directory:

\begin{quote}
\begin{verbatim}
prompt> ant
\end{verbatim}
\end{quote}

\noindent To compile \Chord, run the following command in the same directory:

\begin{quote}
\begin{verbatim}
prompt> ant compile
\end{verbatim}
\end{quote}

\noindent This will compile the following:
\begin{itemize}
\item
the Java source code of \Chord\ from \code{main/src/java/} (and from \code{extra/src/java/} if that directory is present) to Java bytecode in \code{main/classes/}.
\item
the C source code of BDD library \htmladdnormallink{BuDDy}{http://buddy.sourceforge.net/} from \code{main/src/bdd/} to a shared library in \code{main/lib/} (\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and \code{libbuddy.dylib} on MacOS); this library is needed for executing program analyses written in Datalog using \htmladdnormallink{bddbddb}{http://bddbddb.sourceforge.net/} (a BDD-based Datalog solver).
\item
the C++ source code of the Chord instrumentation agent from \code{main/src/agent/} to a shared library in \code{main/lib/} (\code{libchord_instr_agent.so} on all architectures); this agent is needed for computing program scope dynamically and for executing dynamic program analyses.
\end{itemize}

>>>>

You can obtain pre-built binaries as well as the source code of Chord.
The following sections explain both options in detail.

\subsection {Binary Installation}
\label{sec:binary-install}

To obtain Chord's pre-built binaries, download file \xlink{chord-bin-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-bin-2.0.tar.gz}.
It contains the following files:

\begin{enumerate}
\item
\code{chord.jar}, which contains the class files of Chord and of libraries used by Chord.
\item
\code{libbuddy.so}, \code{buddy.dll}, and \code{libbuddy.dylib}: you can keep one of these files
depending upon whether you intend to run Chord on Linux, Windows/Cygwin, or MacOS, respectively.
These files are needed only if you want to use the BuDDy BDD library for executing program
analyses written in Datalog.
\item
\code{libchord_instr_agent.so}: this file is needed only if you want to use the JVMTI-based bytecode
instrumentation agent for executing dynamic program analyses in Chord.
\end{enumerate}

\noindent Novice users can ignore items (2) and (3) above until they become more familiar with Chord.
The provided binaries in those items might not be compatible with your machine, in which case you
will either have to forgo using them (with hardly any noticeable difference in Chord's functionality)
or you will have to build them yourself (see Section \ref{sec:source-install}).

\subsection{Source Installation}
\label{sec:source-install}

To obtain Chord's source code, download the following files:

\begin{enumerate}
\item
Mandatory: file \xlink{chord-src-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-src-2.0.tar.gz}, which contains Chord's source code and the jars of libraries used by Chord.
\item
Optional: file \xlink{chord-libsrc-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-libsrc-2.0.tar.gz}, which contains the source
code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.)
\end{enumerate}

\noindent Alternatively, you can obtain the latest development snapshot from the SVN repository:

\begin{verbatim}
    prompt> svn checkout http://jchord.googlecode.com/svn/trunk/ chord
\end{verbatim}

\noindent Instead of checking out the entire \code{trunk/}, which contains several sub-directories, you can check out
specific sub-directories:
\begin{itemize}
\item \code{main/} contains Chord's source code and the jars of libraries used by Chord.
\item \code{libsrc/} contains the source code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.).
\item \code{test/} contains Chord's regression tests.
\item many more; these might eventually move into \code{main/}.
\end{itemize}

\noindent Files \code{chord-2.0-src.tar.gz} and \code{chord-2.0-libsrc.tar.gz} mentioned above are essentially stable releases of
the \code{main/} and \code{libsrc/} directories, respectively.

### 2011-03-10 main/doc/setting_up.tex
added more documentation

\section{Setting up a Java program}
\label{sec:program-setup}

@@

File \code{chord.properties} specifies properties to be passed to
Chord (alternatively, these properties may be passed on the
command-line, in format \code{-D<name>=<value>}).  A sample such file
for the above program is as follows:

\begin{verbatim}
    chord.main.class=foo.Main
    chord.class.path=classes:lib/jar/taz.jar
    chord.src.path=src:../lib/src
    chord.run.ids=0,1
    chord.args.0="-thread 1 -n 10"
    chord.args.1="-thread 2 -n 50"
\end{verbatim}

Each relative path element in the value of any property named
\code{chord.<...>.path} defined in this file is converted to an
absolute path element with respect to the directory containing this
file, which in the above case is \code{example/}.

\begin{itemize}
\item
\code{chord.main.class} specifies the fully-qualified name of the main
class of the program.
\item
\code{chord.class.path} specifies the application-specific classpath
of the program (the standard library classpath is implicitly
included).
\item
\code{chord.src.path} specifies the Java source path of the program.
All program analyses in Chord operate on Java bytecode.  The only use
of this property is to HTMLize the Java source files of the program so
that the results of program analyses can be reported at the Java
source code level.
\item
\code{chord.run.ids} specifies a list of IDs to identify runs of the
program.  It is used by dynamic program analyses to determine how many
times the program must be run.  An additional use of this property is
to allow specifying the command-line arguments to use in the run
having ID {\tt <id>} via property \code{chord.args.<id>}, as
illustrated by properties \code{chord.args.0} and \code{chord.args.1}
in the above example.
\end{itemize}

@@

\noindent where {\tt <...>} denotes the absolute path of the parent of
directory {\tt example/}.  Property \code{chord.work.dir} specifies
the working directory during Chord's execution.

The above command causes Chord to load the \code{chord.properties}
file if present.  The location of this file may be specified
explicitly via property \code{chord.props.file} on the above command
line.  By default, it is \code{[chord.work.dir]/chord.properties}.

In the above case, Chord does not do much beyond loading the sample
\code{chord.properties} file.  For Chord to do something interesting,
you need to set additional properties, either on the above command
line or in the \code{chord.properties} file, specifying the task(s)
Chord must perform.  All tasks are summarized in Section
\ref{sec:chord-sysprops:task}.  The two most common tasks, described
next, are computing the scope of the given program (Section
\ref{sec:building-scope}) and running analyses on the given program
(Section \ref{sec:running-analysis}).

>>>>

\section{Setting up a Java Program}
\label{sec:setting-up}

@@

@@

\noindent where \code{<...>} denotes the absolute or relative path of the
directory containing the above \code{example/} directory.
Property \code{chord.work.dir} specifies the working directory during Chord's execution.
Chord searches for a file named \code{chord.properties} under that working directory, and
loads upfront all properties defined in that file, if it exists.  Alternatively, these properties
may be defined on the command-line, using the \code{-D<key>=<value>} format.
A sample \code{chord.properties} file for the above program is as follows:

\begin{verbatim}
    chord.main.class=foo.Main
    chord.class.path=classes:lib/jar/taz.jar
    chord.src.path=src:lib/src
    chord.run.ids=0,1
    chord.args.0="-thread 1 -n 10"
    chord.args.1="-thread 2 -n 50"
\end{verbatim}

Each relative file/directory name in the value of any property
defined in this file (e.g., the \code{lib/src} directory name in the value of
property \code{chord.src.path} above) is treated relative to Chord's working directory
(which is specified by property \code{chord.work.dir}).

\begin{itemize}
\item
\code{chord.main.class} specifies the fully-qualified name of the main
class of the program.
\item
\code{chord.class.path} specifies the application classpath
of the program (the standard library classpath is implicitly
included).
\item
\code{chord.src.path} specifies the Java source path of the program.
All program analyses in Chord operate on Java bytecode.  The only use
of this property is to HTMLize the Java source files of the program so
that the results of program analyses can be reported at the Java
source code level.
\item
\code{chord.run.ids} specifies a list of IDs to identify runs of the
program.  It is used by dynamic program analyses to determine how many
times the program must be run.  An additional use of this property is
to allow specifying the command-line arguments to use in the run
having ID {\tt <id>} via property \code{chord.args.<id>}, as
illustrated by properties \code{chord.args.0} and \code{chord.args.1}
in the above example.
\end{itemize}

In the above case, Chord does not do much beyond loading the above
\code{chord.properties} file.  For Chord to do something interesting,
you need to set additional properties, either on the above command
line or in the \code{chord.properties} file, specifying the task(s)
Chord must perform.  All tasks are summarized in Section
\ref{sec:func-props}.  The two most common tasks, described
next, are computing the analysis scope of the given program (Section
\ref{sec:building-scope}) and running analyses on the given program
(Section \ref{sec:running-analysis}).

### 2011-03-10 main/doc/whatis_chord.tex
added more documentation

Chord is a program analysis platform that enables 
productively design, implement, evaluate, and combine a
broad variety of program analyses for Java bytecode. It has the
following key features:

\begin{itemize}
\item
{\bf Stand-alone:} It provides various off-the-shelf analyses (e.g., various
call-graph analyses; points-to analyses; thread-escape analysis;
static slicing; static and dynamic concurrency analyses for finding
races, deadlocks, and atomicity violations; etc.)

\item
{\bf Flexible:} It allows users to express a broad range of analyses,
including both static and dynamic analyses, analyses written
imperatively in Java or declaratively in Datalog, summary-based as
well as cloning-based inter-procedural context-sensitive analyses,
iterative refinement-based analyses, client-driven analyses, and
combined static/dynamic analyses.

\item
{\bf Efficient:} It executes analyses in a demand-driven fashion, caches
results computed by each analysis for reuse by other analyses without
re-computation, and can execute analyses without dependencies between
them in parallel.

\item
{\bf Deterministic:} It guarantees that the result is the same regardless of
the order in which different analyses are executed; moreover, results
can be shared across different runs.
\end{itemize}

>>>>

Chord is a program analysis platform that enables users to
productively design, implement, evaluate, and combine a
broad variety of program analyses for Java bytecode. It has the
following key features:

\begin{itemize}
\item
{\bf Stand-alone:} It provides various off-the-shelf analyses (e.g., various
call-graph analyses; points-to analyses; thread-escape analysis;
static slicing; static and dynamic concurrency analyses for finding
races, deadlocks, and atomicity violations; etc.)
\item
{\bf Flexible:} It allows users to express a broad range of analyses,
including both static and dynamic analyses, analyses written
imperatively in Java or declaratively in Datalog, summary-based as
well as cloning-based inter-procedural context-sensitive analyses,
iterative refinement-based analyses, client-driven analyses, and
combined static/dynamic analyses.
\item
{\bf Efficient:} It executes analyses in a demand-driven fashion, caches
results computed by each analysis for reuse by other analyses without
re-computation, and can execute analyses without dependencies between
them in parallel.
\item
{\bf Deterministic:} It guarantees that the result is the same regardless of
the order in which different analyses are executed; moreover, results
can be shared across different runs.
\end{itemize}

### 2011-03-10 main/doc/writing_analysis.tex
added more documentation

\section{Writing a Program Analysis}
\label{sec:writing-analysis}

TODO

>>>>

\section{Writing an Analysis}
\label{sec:writing-analysis}

Chord is built around the notion of modeling a program with relations.
A Chord analysis is a program component that operates on these relations.
It may consume zero or more relations, do some work on them, and then output zero or more relations.
It may also print output in some other format, perhaps human-readable text or HTML.
``Relation'' here has the same meaning it does in the context of relational databases: a set of tuples, containing elements of one or more domains.  
%For example, the relation \texttt{reachableM(m:M)} identifies a subset of

An analysis can be written in one of two ways: either as a Java class or as a datalog program.

\subsection{Dependency information}

Chord has two kinds of projects: ``Classic'' and ``Modern''.  Despite the names, classic is effectively the standard at this point.
In the classic model, each analysis specifies the domains/relations it consumes and that it produces.
This defines an implicit dependency graph between analyses.  For a datalog analysis, the ``input'' and ``output'' keywords on relations specify the requirements and results of an analysis.  For a Java analysis, the \texttt{@Chord} annotation does so.

\subsection{Domains}

In Chord, a domain is represented as a list of possible values. 
The value types must be comparable and hashable.
Strings work, as do integers, \texttt{joeq} quads, fields, types, and so forth. 
%
% What are the required types?

Both domains and relations have names. The built-in relations are usually identified by a single capital letter.
Chord by default includes code for constructing a number of domains.

It is sometimes useful to construct new domains in the middle of an analysis, perhaps as a way of compactly describing some entities that required nontrivial work to find. Domains, for now, must be defined in Java. One key requirement is that construction must be deterministic: if the elements in a domain can be reordered from execution to execution on the same program, this can result in subtle and hard-to-track-down bugs.

\begin{itemize}
\item
B is the domain of basic blocks.
\item
C is the domain of abstract calling contexts.
\item
E is the domain of program points that read or write a field.
\item
F is the domain of fields inside a class.
\item
H is the domain of allocation sites.
\item
I is the domain of invocation statements (method calls)
\item
L is the domain of lock acquisition points.
\item
M is the domain of program methods.
\item
P is the domain of simple statements. %SAY MORE HERE?
\item
R is the domain of lock release points.
\item
T is the domain of types (classes and primitives).
\item
V is the domain of reference-typed variables.
\item
W is the domain of loop head statements. %SAY MORE HERE?
\item
Z is a domain of integers corresponding to argument positions.The size of the domain is chosen to 
accommodate the method with the greatest number of arguments.
\item
ZZ is another domain of integers, this one of fixed size and available for other uses.
\end{itemize}

\subsection{Core relations}

Chord includes code to build a very large number of relations. These standard relations
are very useful in writing your own analyses.
For example, relation \texttt{sub} includes all pairs of types \texttt{(t1, t2)} where \texttt{t2} is a subclass of \texttt{t1}.
Most of these standard relations are in the package \texttt{chord.rels}.

\subsection{Java analyses}

A Java analysis should extend JavaAnalysis or ProgramRel.  It's usually easiest if an analysis
only produces one relation, in which case extending ProgramRel is the best approach.
If you extend ProgramRel, all the work of creating and saving the relation is taken care of for you;
all you need to do is add tuples to the relation.Typically you do this by calling
\texttt{super.add(...)} with the tuple you wish to add to the relation.
For forther convenience, you can implement one of the Visitor interfaces defined in 
\texttt{chord.program.visitors}.  This will force you to implement some \texttt{visit} methods that iterate over
all elements in some domain, such as all methods or all invocation points.

Sometimes, you need to create two different relations at the same time. In this case, the right approach is to
extend JavaAnalysis, and explicitly create and save the relations on your own.
\textbf{Should reference JavaDoc.  Should make sure JavaDoc explains what you need to do to fill and save a relation}.

Regardless of which of these approaches you take, every Java analysis must have an \texttt{@Chord} annotation.
This annotation is a set of named fields.
At a minimum, this annotation must specify the name of the analysis.
If you extend ProgramRel, the name of the analysis will be treated as the same as that of the generated relation.
For example, \texttt{@Chord(name = "F")} is a valid annotation. 
The name of the Java class that generates the relation is ignored by Chord -- only the annotation matters.

\textbf{MAYUR, IS SIGN MANDATORY?   IF NOT, WHEN CAN IT BE OMITTED?}

There are also other annotation fields available.
You may specify the dependencies of your analysis via \texttt{consumes = \{ "xyz"\}}.

If your analysis does not extend ProgramRel, you should also specify the relations that your 
analysis produces. You do this via the \texttt{produces} relation.

%Should mention the signs fields.
%
%

\noindent  The default value of property \code{chord.java.analysis.path}
is \code{chord.std.java.analysis.path}:\code{chord.ext.java.analysis.path},
where the defalt value of \code{chord.std.java.analysis.path} is the location of file \code{chord.jar}, and
the default value of \code{chord.ext.java.analysis.path} is empty (users defining their own Java analyses are expected to
set the latter property).

\noindent  The default value of property \code{chord.dlog.analysis.path}
is \code{chord.std.dlog.analysis.path}:\code{chord.ext.java.analysis.path},
where the defalt value of \code{chord.std.java.analysis.path} is the location of file \code{chord.jar}, and
the default value of \code{chord.ext.java.analysis.path} is empty (users defining their own Java analyses are expected to
set the latter property).

### 2011-03-10 main/doc/writing_datalog_analysis.tex
added more documentation

\section{Writing a Datalog analysis}
\label{sec:writing-datalog-analysis}

@@

Any line that begins with a {\tt \#} is regarded a comment, except a
line of the form {\tt \#} {\tt name=<...>}, which specifies the name
{\tt <...>} of the Datalog analysis.
Each Datalog analysis is expected to have exactly one such line.
The above Datalog analysis is named {\tt datarace-dlog}.
In Chord, all Datalog analysis names have suffix
{\tt -dlog} and all Java analysis names have suffix {\tt -java}, but
this is merely a convention that user-defined analyses are free to deviate from.
The name of each analysis, written in Datalog or in Java, is expected to be
unique across all analyses in scope (i.e., across all analyses in paths
specified by properties \code{chord.dlog.analysis.path} and \code{chord.java.analysis.path}).

>>>>

\section{Writing a Datalog Analysis}
\label{sec:writing-datalog-analysis}

@@

Any line that begins with a {\tt \#} is regarded a comment, except a
line of the form ``{\tt \#} {\tt name=<...>}", which specifies the name
{\tt <...>} of the Datalog analysis.
Each Datalog analysis is expected to have exactly one such line.
The above Datalog analysis is named {\tt datarace-dlog}.
In Chord, all Datalog analysis names have suffix
{\tt -dlog} and all Java analysis names have suffix {\tt -java}, but
this is merely a convention that analysis writers are free to deviate from.
The name of each analysis, written in Datalog or in Java, is expected to be
unique across all analyses in scope (i.e., across all analyses in paths
specified by properties \code{chord.dlog.analysis.path} and \code{chord.java.analysis.path}).
